Modelling the Retinotectal projection - can competition provide a stopping mechanism?

### 2021-05-12 paper/paper.tex
Model tweaked to have retina and tectum with receptors and ligands on each

The retinotectal projection has proved to be a deep mine of information for
the study of how the cells of the central nervous system are accurately
connected together into functional networks. This projection connects the
light-gathering cells in the retina to movement related cells in the optic
tectum (known as the superior colliculus in mammals). Light sources
originating close to each other in the environment tend to activate retinal
cells situated close together in the eye so that an image of the environment
is formed on the retinal surface. It has been discovered that the topography
of the retina is preserved within brain regions that process this information
such that cells which are adjacent within the retina primarily excite cells
adjacent in the tectum. This indicates that during development there must
exist a mechanism which ensures the correct wiring of the axons which leave
the retina and connect to cells in the tectum.

One reason for the success of the study of the retinotectal projection is its
capacity to be experimentally manipulated. In some non-mammalian species, both
the retina and the tectum can be partially ablated, or even physically
reorganised in-vivo, after which axons regrow to restore the order and
function of the system for the individual animal. This manipulability was
exploited in influential work by R. W. Sperry and co-workers during the
mid-twentieth century, leading to Sperry's 1963 summary of
the \emph{chemoaffinity theory}~\cite{sperry_chemoaffinity_1963} which
proposes the existence of morphogenetic gradients that guide axons to their
destination. The chemoaffinity theory was given robust support by the
discovery of the ephrin ligands and their receptors~\cite{first_ephrin_paper}
which have been shown to form into graded expression fields in the
retina~\cite{ephrin_retina}, tectum~\cite{ephrin_tectum} as well as in other
sensory systems, such as the somatosensory
system~\cite{vanderhaegen_as_in_elife_paper}. The Ephrin ligands have a clear
effect on axonal outgrowth, as shown in in-vitro~\cite{one} and
in-vivo~\cite{another} studies.

@@

>>>>

The retinotectal projection has proved to be a deep mine of information for
the study of how the cells of the central nervous system are accurately
connected together into functional networks. This projection connects the
light-gathering cells in the retina to movement-related cells in the optic
tectum (known as the superior colliculus in mammals). Light sources
originating close to each other in the environment tend to activate retinal
cells situated close together in the eye so that an image of the environment
is formed on the retinal surface. It has been discovered that the topography
of the retina is preserved within brain regions that process this information
such that cells which are adjacent within the retina primarily excite cells
adjacent in the tectum. This indicates that during development there must
exist a mechanism which ensures the correct wiring of the axons which leave
the retina and connect to cells in the tectum.

One reason for the success of the study of the retinotectal projection is its
capacity to be experimentally manipulated. In some non-mammalian species, both
the retina and the tectum can be partially ablated, or even physically
reorganised in-vivo, after which axons regrow to restore the order and
function of the system for the individual animal. This manipulability was
exploited in influential work by R. W. Sperry and co-workers during the
mid-twentieth century, leading to Sperry's 1963 summary of
the \emph{chemoaffinity theory}~\cite{sperry_chemoaffinity_1963} which
proposes the existence of morphogenetic gradients that guide axons to their
destination. The chemoaffinity theory was given robust support by the
discovery of the ephrin ligands and their receptors~\cite{first_ephrin_paper}
which have been shown to form into graded expression fields in the
retina~\cite{ephrin_retina}, tectum~\cite{ephrin_tectum} as well as in other
sensory systems, such as the somatosensory
system~\cite{vanderhaegen_as_in_elife_paper}. The Ephrin ligands have a clear
effect on axonal outgrowth, as shown in in-vitro~\cite{one} and
in-vivo~\cite{another} studies.
%
% A-P:
%
% EphA and ephrinAs: Anteroposterior axis of retiontectal projection
% (Nasal-Temporal on retina; Anteroposterior on tectum).
%
% RGM (tectum) Neogenin (retina)
%
% D-V:
%
% EphB and ephrin B: Dorsoventral on retina; dorsoventral on tectum.
%
% En-2: Repels temporal growth cones of xenopus ret-tec axons and attracts
% nasal ones. Brunet et all, Nature, 2005
%
% Ryk/Fz (retina) and Wnt3 (tectum) (Flanahan, 2006)
%

@@

New introduction: Simpson and Goodhill presented a model in which the
chemoaffinity mechanism is non-biological. We offer a model in which a
competition mechanism counteracts the drive of axons along the ephrin
gradient. Will we need the axon-axon interaction mechanism based on
non-same-Eph repulsion?
